---
title: "Illustrating input of fixed width format files"
author: "Steve Simon"
date: "Sunday, June 12, 2016"
output: html_document
---

You should generally use delimited format files (e.g., comma separated
values) when they are available, as they generally read in easier and
with less fuss. Some data files, however, are not delimited, or the
delimiters aren't consistent. In these settings, you would have to
read the date using a fixed width format.

The first example of reading in a fixed width format comes from
the Journal of Statistics Education dataset archive page. It is
a salary survey of college professors conducted by the American
Association of University Professors (AAUP).

Here's a description of the structure of the data set.

VARIABLE DESCRIPTIONS (AAUP2.DAT)

Fixed column format with two data lines per school

Line #1

 1 -  5   FICE (Federal ID number)

 7 - 37   College name

38 - 39   State (postal code)

40 - 43   Type  (I, IIA, or IIB)

44 - 48   Average salary - full professors

49 - 52   Average salary - associate professors

53 - 56   Average salary - assistant professors

57 - 60   Average salary - all ranks

61 - 65   Average compensation - full professors

66 - 69   Average compensation - associate professors

70 - 73   Average compensation - assistant professors

74 - 78   Average compensation - all ranks

Line #2

 1 -  4   Number of full professors

 5 -  8   Number of associate professors

 9 - 12   Number of assistant professors
 
13 - 16   Number of instructors

17 - 21   Number of faculty - all ranks

```{r import-from-web}
help("read.fwf")
ends_line1 <- c(5, 6, 37, 39, 43, 48, 52, 56, 60, 65, 69, 73, 78)
ends_line2 <- c(4, 8, 12, 16, 21)
widths_line1 <- diff(c(0,ends_line1))
widths_line2 <- diff(c(0,ends_line2))
fn <- "http://www.amstat.org/publications/jse/datasets/aaup2.dat.txt"
aaup <- read.fwf(fn, c(widths_line1,widths_line2), stringsAsFactors=FALSE)
head(aaup)
str(aaup)
```

Notice that some of the variables that look like they should be numeric
are actually stored as character. Usually this is caused by a few stray
values which are not numeric. Here's a quick way to list the 
non-numeric varialbes.

```{r list_non-numeric_values}
non_numeric_list <- which(is.na(as.numeric(aaup$V6)))
table(aaup$V5[non_numeric_list])
```

The non-numeric code in this variable is probably a missing value code.

